---
title: "Using mScatter"
author: "Data Computing"
date: "May 27, 2015"
output: ioslides_presentation
---

## Steps for making a graphic

1. Pick a modality
    - Scatter, density, bar, map
2. Choose variables for the frame
    - maps: frame is predefined
    - densities: y is pre-defined
    - facets can represent a third variable
3. Map variables to aesthetics
    - color, shape, length, 
    - x & y were mapped when choosing the frame
4. Refine as needed
    - x & y axis labels
    - many others
    
## mScatter et al.

These simply provide a mouse-based interface for steps two and three.

* It's still you who has to make the decisions.  
    `mScatter()` et al. let you evaluate possible choices quickly.



### Tips

* Take a "small" subset of your data to develop the graphic, e.g. 1000 points.  This will speed up your explorations.
* When you're finished, you can copy-and-paste the graphic, but better to ...
* Copy-and-paste the `ggplot2` **command** generated from your choices.
* You might prefer `mPlot()` from the `mosaic` package.

## Exercise

Make this (awful) graphic with `mScatter()` using the `NHANES::NHANES` data. (You may need to install `NHANES` from CRAN.)

```{r include=FALSE}
require(NHANES)
require(DCFinteractive)
```

```{r echo=FALSE}
require(NHANES)
data(NHANES, package = "NHANES")
require(DCFinteractive)
```

```{r eval=FALSE, echo=FALSE}
DCFinteractive::mScatter(NHANES)
```

```{r echo=FALSE, warning=FALSE}
ggplot(data=NHANES,aes(x=Age,y=BMI))+geom_point()+aes(colour=Depressed)+facet_wrap(~Smoke100n,ncol=4) 
```



    
    