---
title: "Explaining Shiny (Brief Edition)"
author: "Daniel Kaplan"
date: "May 2015"
output: 
  html_document:
    toc: FALSE
    number_sections: FALSE
runtime: shiny
---

```{r include=FALSE}
library(shiny, quietly=TRUE)
library(DCF, quietly=TRUE)
knitr::opts_chunk$set(tidy = FALSE)
```

This is an RMarkdown ⇒ HTML document that has interactive controls.  It's written with the Shiny system.

An interactive document has *user inputs* and *display outputs*. The user sets the inputs as desired, the outputs respond accordingly. In writing an interactive document, you have to specify four things:

1. What form for the various inputs: sliders, radio buttons, etc.
#. What form for the outputs: graphics, tables, etc.
#. A session with the data tables, packages, and all the other ordinary R objects that you will want to use
#. Programming logic that transforms inputs to output



## (1) Input Widgets

To create an input widget, you write a chunk in which you call the appropriate `shiny` function. The widget will be displayed as the output of the chunk. For instance:

```{r}
sliderInput(inputId = "nbins",
            label = "Bin width",
            min = 0.1, max = 2, value = 0.5, step = 0.1)
```

In practice, you would suppress display of the command using `echo = FALSE` as a chunk argument.  But this is a document about how to use Shiny; the commands are part of the explanation.

## (2) Display Outputs

Just like the input widgets, the output displays are created by `shiny` functions.  There are many sorts: `renderPlot()`, `renderDataTable()`, `renderText()` and so on.  Here's a plot display:

```{r}
renderPlot(make_my_plot())
```

"What's the `make_my_plot()` function?" you ask.  This will contain the programming logic for filling in the display. Read on.

## (3) An R Session

This is where you will create data frames, load packages, and create all those regular R things that you'll need for your display. In this example, we'll use data from the `mosaicData` package.

```{r}
library(mosaicData)
```

## (4) Programming Logic

*Programming logic* refers to the transformation of the information from the *input widgets* and the *R session* into the values that will be displayed. For this example, that logic will be contained within the `make_my_histogram()` function. Here's a simple implementation.

```{r}
make_my_plot <- reactive( 
  {
    vals <- data.frame(x = rnorm(1000))
    ggplot(vals, aes(x = x)) + 
      geom_histogram(binwidth = input$nbins)
  }
)
```

In many ways, `make_my_plot()` is an ordinary function. The statements in the function are collected in `{` and `}` brackets, in the normal way. `make_my_plot()` invokes `ggplot()` to generate graphics. The return value is the graphical object created by `ggplot()`, just as it would be in a normal function.  But there are some subtle differences that you should note.

* `make_my_plot()` is created by the `reactive()` function rather than using the ordinary `function()` syntax.  `reactive()` is a special way of creating functions that knows how to communicate with the input widgets. In particular, any time the input widgets named in `make_my_plot()` change, the function will be re-evaluated. And, any statements that contain a call to `make_my_plot()` will also be re-evaluated.  This includes the `renderPlot()` statement used to define the display output.
* Within `make_my_plot()`, the connection to the input widgets is made by grabbing their value with the expression `input$nbins`.  Recall that `"nbins"` is the `inputId` for the slider.  If you have other input widgets, you can refer to them as `input$`*widget_name*.
* Unlike the way `function()` is used, the collection brackets `{` and `}` are *inside* the parentheses of `reactive()`.  That is, the familiar form is 

`function(` *args* `) {` *statements* `}` 

but for the reactive form^[Why did [Joe Cheng](https://www.linkedin.com/in/jcheng), the inventor of Shiny, write things this way?  He had little choice.  `function()` isn't actually a function in R.  It is a *special form* that is part of the core R specification.  As such, `function()` gets to do special things that contributed functions like `reactive()` are not allowed to do.] it is

`reactive( {` *statements* `} )`

It's easy to be confused by this.

# Build your own app!

You can build it right here, leaving the example statements in place.  Just make sure to use unique names for any widgets and reactive functions that you create.

## Some ideas:

### Restaurant Inspections

Data on restaurant inspections in New York City is provided by the `RestaurantsNYC` package.  If you haven't done so already, you will need to install the package:
```{r eval=FALSE}
devtools::install_github("dtkaplan/Data/RestaurantsNYC")
```

Write a new histogram display app like the one above but customize it for the scores given to restaurants on each inspection.  Display a histogram of scores arranged so that you can interactively change the bin size.  For small bins, a feature emerges that isn't evident when using just, say, 5 or 10 bins.

Elaboration:

* But in a slider to choose a time interval in days, with a default set to about 30 days.  Wrangle the inspection data so that you can compare density plots of the scores from inspections that happen within the selected interval after the previous inspection and those that come don't.

### Baby Names

Using the `BabyNames` data table, implement your own version of [this app](http://www.ssa.gov/oact/babynames/#ht=2) from the Social Security Administration.  Note: You will **not** need the "Go" or "Reset" buttons.
 