---
title: "USCOTS data science workshop: intro to text mining"
author: "Anonymous until you change this!"
date: ""
output: 
  html_document:
    fig_height: 3
    fig_width: 5
  pdf_document:
    fig_height: 3
    fig_width: 5
  word_document:
    fig_height: 3
    fig_width: 5
---


```{r include=FALSE}
# Don't delete this chunk if you are using the mosaic package
# This loads the mosaic and dplyr packages
require(mosaic)
require(DCF)
```

```{r include=FALSE}
# Some customization.  You can alter or delete as desired (if you know what you are doing).

# This changes the default colors in lattice plots.
trellis.par.set(theme=theme.mosaic())  

# knitr settings to control how R chunks work.
require(knitr)
opts_chunk$set(
  tidy=FALSE,     # display code as typed
  size="small"    # slightly smaller font for code
)
# This loads the mosaic data sets.  (Could be deleted if you are not using them.)
require(mosaicData)                
```

### Introduction

The `aRxiv` package facilitates access to http://arXiv.org, a repository
of electronic preprints for a number of scientific disciplines.  It receives
many thousands of new submissions each month.  In this markdown file, we will explore some 
analyses of the text in some of these abstracts.

To avoid overloading the arXiv server, we will use data that have been previously downloaded using the following commands:
```{r eval=FALSE}
require(aRxiv)
efron <- arxiv_search(query='cat:stat*', limit=1000)
names(efron)
dim(efron)
efron <- mutate(efron, submityear =
  year(sapply(str_split(submitted, " "), "[[", 1)))
with(efron, table(submityear))
```






This file can be knit to HTML, PDF, or Word.  In RStudio, just select the desired output file type
and click on `Knit HTML`, `Knit PDF`, or `Knit Word`.  Use the dropdown menu next to that to 
change the desired file type.