---
title: "Bike Sharing Basics"
author: "Data Computing"
date: "USCOTS 2015"
output: html_document
---

```{r include=FALSE}
library(DCF)
library(ggplot2)
library(mosaic)
```


```{r cache=TRUE}
Stations <- mosaic::read.file("http://tiny.cc/dcf/DC-Stations.csv")
data_site <- "http://tiny.cc/dcf/2014-Q4-Trips-History-Data-Small.rds"
Trips <- readRDS(gzcon(url(data_site)))
```


#### **1.** Which station has the most outgoing rentals?

```{r}
Trips %>%
  group_by(sstation) %>%
  summarise(rental_volume = n()) %>%
  arrange(desc(rental_volume)) %>%
  head(1)
```

#### **2.** Which station has the most outgoing rentals on Saturday?

```{r}
Trips %>% 
  mutate(day_of_week = lubridate::wday(sdate)) %>%
  filter(day_of_week == 7) %>%
  group_by(sstation) %>%
  summarise(rental_volume = n()) %>%
  arrange(desc(rental_volume)) %>%
  head(1)
```

#### **3.** Which origin/destination pair of stations has the most traffic?

```{r}
Trips %>%
  group_by(sstation, estation) %>%
  summarise(rental_volume = n()) %>%
  arrange(desc(rental_volume)) %>%
  head(1)
```

#### **4.** Is mean daily outbound rental volume different on weekdays and weekends taking all stations together in aggregate. 

```{r}
Trips %>% 
  mutate(day_of_week = lubridate::wday(sdate),
         weekend = day_of_week %in% c(1,7)) %>%
  group_by(weekend, day_of_week) %>%
  summarise(rental_volume = n()) %>%
  summarise(daily_mean = mean(rental_volume, na.rm=TRUE)) 
```

    
#### **5.** (Optional) Same as (4), but give a separate answer for each station.

```{r}
Trips %>% 
  mutate(day_of_week = lubridate::wday(sdate),
         weekend = day_of_week %in% c(1,7)) %>%
  group_by(sstation, weekend, day_of_week) %>%
  summarise(rental_volume = n()) %>%
  summarise(daily_volume = mean(rental_volume, na.rm=TRUE)) %>%
  spread(key = weekend, value = daily_volume) %>%
  head() # just to avoid printing all the results here.
```

A subtlety: Note that each `summarise()` statement *removes* one of the grouping variables.  The others survive.  You can see what groups are present at any point with the `groups()` statement.